### 2017-04-05 paper.tex
Half-rewritten into high-to-low order

\documentclass[twocolumn]{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}
\title{Factoring with $n+1$ clean qubits and $n$ dirty qubits}
\author{Craig Gidney}

\begin{document}
\maketitle

\begin{abstract}
\em
We present size-efficient reversible classical circuits for performing various arithmetic operations aided by ancilla in an unknown state.
We improve the number of qubits needed to factor an $n$-bit number from $2n+2$ clean qubits \cite{takahashi2006, haner2016} to $n+1$ clean qubits and $n$ dirty qubits, without increasing the asymptotic size or depth of the circuit.
\end{abstract}

\section{Introduction}

When constructing quantum circuits, or classical reversible circuits, the number of available ancilla is an important resource.
An ancilla is an extra bit or qubit available for use by a circuit as temporary workspace.
Ancilla may be initialized to a known state (``clean bit"), or be given to the circuit in an unknown state (``dirty bit") that must be restored before the circuit finishes.
Clean bits are more valuable, allowing for simpler and more compact circuit constructions, but dirty bits are more plentiful, since any temporarily unused bit is a borrowable dirty bit.

Because one part of a circuit can borrow dirty bits from another part of the same circuit, circuit constructions that require only dirty bits are easier to apply under tight space constraints, or on circuit topologies where other ancilla are too far away to be acquired quickly.
For example, in \cite{haner2016} H\"{a}ner et al. used dirty bits to improve the cost of performing Shor's algorithm, when limited to using only $2n+2$ qubits and Toffoli based circuits, from ??????? to $O(n^3)$ depth and $O(n^3 \lg n)$ size.
The construction in this paper improve on \cite{haner2016} by cutting the number of required qubits to factor an $n$-bit number from $2n+2$ to $2n+1$.

Our paper is structured as follows.
The introduction describes the various sections and also itself for no discernible reason.
Section \ref{sec:review} discusses existing circuit constructions for various arithmetic tasks.
In section \ref{sec:construct} we iteratively reduce period finding into simpler and simpler arithmetic operations, until we reach constant-sized gates, while tracking the number of required dirty ancilla.
Then, in section \ref{sec:costs}, we discuss the novelty and comparative costs of the presented arithmetic circuit constructions.
Finally, \ref{sec:conclusions} concludes and summarizes.

All circuit daigrams assume a 2s-complement representation of integers, and have the least significant bit towards the top.
For clarity, circuit diagrams will divide compound operations into separate input and output parts, with inputs shaded gray.

\section{Existing Constructions} \label{sec:review}

\subsection{Addition}

In \cite{draper2000}, Draper describes an addition circuit that uses the quantum Fourier transform to convert controlled phasing into addition.
Draper notes that the construction uses $O(N)$ depth and $O(N^2)$ controlled rotations, but approximation techniques \cite{barenco1996} \cite{cleve2000} can improve the depth and size to $O(\lg N)$ and $O(N \lg N)$ respectively.

Both \cite{cuccaro2004} and \cite{van2004} describe classical reversible ripple-carry adders that use a single clean ancilla, $O(N)$ depth, and $O(N)$ size.
Takahashi \cite{takahashi2005} showed how to use the most-significant-bit (MSB) of the input register, instead of a separate ancilla, to propagate the carry signal.
Takashi later showed how to trade depth against size \cite{takahashi2009}.

We generalize the ideas in \cite{takahashi2005} to apply to additions where the input and target registers have differing sizes.

\subsection{Offset}

An offset gate adds a compile-time constant into a register.
Some reversible adder constructions can be used as offset gates, but others can't because they use qubits in the input register as temporary workspace.
An offset gate has no input register.

Draper's QFT-based addition construction \cite{draper2000} doesn't use the input register as workspace, and so trivially generalizes into an offset gate.

\cite{haner2016} describes a Toffoli-based divide-and-conquer offset circuit that requires $O(N)$ depth, $O(N \lg N)$ size, and $n/2$ dirty ancilla.
[[[ Actually, only 1 is required by their construction. 0 if you go quantum. And when I emailed them about it they said they would fix the oversight. So I'm not sure how to describe that. It is important for this paper that we have an non-modular offset circuit that only needs 1 dirty ancilla. ]]]

We describe a circuit for the special case of offset-by-1, i.e. incrementation.
Our increment circuit uses $O(N)$ depth, $O(N)$ size, one dirty ancilla, and no non-classical operations.
We also note how to cut the ancilla with quantum operations.
This construction is a simplified version of our unpublished incrementer described in \cite{gidney2015}.

\subsection{Comparison}

Comparison operators, which toggle a target bit based on if one input is larger, smaller, and/or equal to another, can be implemented efficiently using adders \cite{takahashi2005}.

Due to generalizing Takahashi's adder \cite{takahashi2005} to apply to different sizes, we also generalize Takahashi's comparators to apply to registers of different sizes.

\subsection{Modular Addition}

\cite{vedral1995} describes a modular adder that uses $O(N)$ clean ancilla.
It temporarily stores a comparison in a clean ancilla, performs a normal addition, uses a subtraction conditioned on the ancilla to wrap the result back into range, then performs a cleanup comparison to clear the ancilla.

All but one of the $O(N)$ clean ancilla used by \cite{vedral1995} are due to using inefficient addition circuits.
Using later zero-ancilla adder constructions reduces xthe number of clean ancilla required to perform a modular addition to just one.

Our modular adder construction also requires a single ancilla, but the ancilla is no longer required to be clean.

\subsection{Period Finding}

...

\section{From Period Finding to Toffolis, using $2n+1$ qubits} \label{sec:first-circuit} \label{sec:construct}

\subsection{Period Finding}

We begin with the abstract period finding circuit from ???, shown in figure \ref{fig:period-finding-solo-phase-qubit}.
This circuit uses (a) controlled modular multiplication, (b) measurement, (c) X-axis rotations parametrized by previous measurements, and (d) qubit resets.
The only non-trivial operation is (a), controlled modular multiplication an $m$ qubit register, for which we will need $m$ clean ancilla.

\begin{figure}
  \centering
  \makebox[\linewidth]{
    \includegraphics[width=\linewidth]{assets/shor-period-finding.png}
  }
  \caption{\em
	High-level period finding circuit [[[ cite ]]].
    The triangle indicates how many ancilla are needed, by our constructions, to perform the linked operation.
  }
  \label{fig:period-finding}
\end{figure}

\begin{figure*}
  \centering
  \makebox[\linewidth]{
    \includegraphics[width=\linewidth]{assets/shor-period-finding-solo-phase-qubit.png}
  }
  \caption{\em
	Period finding with a single phase-estimation qubit [[[ cite ]]].
	The small $\tiny \oplus$ are ``X-axis controls".
	An X-axis control is equivalent to a normal control, but with a Hadamard gate applied before and after.
	It conditions on the state $\frac{1}{\sqrt 2}|0\rangle - \frac{1}{\sqrt 2}|1\rangle$ instead of on the state $|1\rangle$.
  }
  \label{fig:period-finding-solo-phase-qubit}
\end{figure*}

\subsection{Controlled Modular Multiply}

We do controlled modular multiplication with the same technique used in ???: multiply-accumulate into an extra register, then erase the input with an inverse multiply-accumulate.
See figure \ref{fig:controlled-modular-multiply}.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/controlled-modular-multiply.png}
  \caption{\em
    Reducing a controlled modular multiplication into two controlled modular multiply-accumulates.
    Based on ???.
  }
  \label{fig:controlled-modular-multiply}
\end{figure}

The controlled modular multiply-accumulates require no ancilla.

\subsection{Controlled Modular Multiply-Accumulate}

To perform multiply-accumulation, we use a shift-and-add algorithm.
We right-shift (i.e. divide by 2) the target register $m-1 {\pmod R}$ times, then begin iteratively left-shifting (i.e. multiplying by 2) and conditionally adding $K$ into the target based on the appropriate input bit.
See figure \ref{fig:controlled-modular-multiply-accumulate}.

Both the modular bit-shift and conditional offset operations require a single dirty bit, but have more than enough available to borrow.
Also, both operations will be implemented in terms of what we call a pivot flip operation.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/controlled-modular-multiply-accumulate.png}
  \caption{\em
    Reducing a controlled modular multiply-accumulate into multiplications by 2 and controlled modular additions.
    Note that, although the controlled modular offset operations require a dirty ancilla, they can borrow it from one of the wires they don't affect.
    Based on ???.
  }
  \label{fig:controlled-modular-multiply-accumulate}
\end{figure}

\subsection{Pivot-Flip}

A "pivot-flip" is an operation that reverses the order of states less than a given pivot value, without affecting other states.
For example, a pivot-flip with the pivot equal to 4 would swap $|0\rangle$ and $|3\rangle$, swap $|1\rangle$ and $|2\rangle$, and leave all other states untouched.
Pivot-flips are useful for implementing cyclic operations, such as modular addition.

The exact permutation performed by a pivot-flip with pivot equal to $k$ is:

$$\text{PivotFlip}_k = \sum_{i=0}^{k-1} |k-i-1\rangle \langle i| + \sum_{i=k}^{N-1} |i\rangle \langle i|$$

To perform a pivot-flip efficiently, we use the fact that $x \rightarrow \lnot(x - a)$ nearly does what is required.
As shown in figure \ref{fig:double-flip}, the permutation $x \rightarrow \lnot(x - a)$ flips the range below $a$ but also flips the range above-and-including $a$.
However, because $x \rightarrow \lnot(x - a)$ is its own inverse and doesn't change whether any value is less than the pivot, the operation can be controlled by a dirty ancilla being toggled by a comparison against $a$.

See figure \ref{fig:const-pivot-flip} for the construction (as well as a less-than-or-equal-to-pivot variant).
For states that are less than the pivot $a$, the dirty ancilla is being toggled so exactly one of the controlled flips will happen.
For other states, the ancilla does not change and so the controlled flip happens either no times or two times (undoing itself) for a net effect of no effect.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/double-flip.png}
  \caption{\em Subtraction followed by logical negation performs flips both above and below $k$.}
  \label{fig:double-flip}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/pivot-flip.png}
  \caption{\em Pivot flip circuits.
  When the top wire bundle is a register, the subtractions are controlled by adding a control to any CNOTs, increments, and decrements in figures \ref{fig:inlineadder} or \ref{fig:inline-adder-into-large}.
  When the top wire bundle is a compile-time constant, the subtractions use the construction from \cite{haner2016} and are controlled as shown in figure \ref{fig:controlled-offset}.}
  \label{fig:const-pivot-flip}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/controlled-offset.png}
  \caption{\em Offset circuit from \cite{haner2016}, modified to include a control.}
  \label{fig:controlled-offset}
\end{figure}

The cost of a $\text{PivotFlip}$ depends on if the pivot is a compile-time constant or not.
When the pivot is input from another register, the $O(N)$ inline adder can be used.
But when the pivot is a constant, the $O(N \lg N)$ size and $O(N)$ depth offset gate from \cite{haner2016} is used instead.

\subsection{Modular Doubling}

If the modulus $R$ is even, modular doubling is irreversible and therefore not unitary (it can be performed, but it must create garbage bits leading to decoherence).
So we assume that the modulus $R$ is odd.

To multiply a register by 2, we use a pivot-flip operation, a subtraction, and CNOTs.
See \ref{fig:modular-double}.
This construction requires a dirty ancilla because the pivot-flip operation covers all wires but requires a dirty ancilla.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/modular-double.png}
  \caption{\em
    Modular doubling.
    Assumes $R$ is odd.
    Requires 1 dirty ancilla.
    Trashes out-of-range values (ones greater or equal to $R$).
  }
  \label{fig:modular-double}
\end{figure}

https://arxiv.org/abs/1202.6614

https://arxiv.org/abs/1110.2574

[[[diagram of their construction]]]

[[[diagram of their doubling]]]

The construction requires $n+1$ clean ancilla.
$n$ for the second register, and $1$ for holding a carry bit during the doubling step.
By using the doubling construction in figure [[[whatever]]], the extra ancilla can be dirty instead of clean.

[[[diagram of my doubling]]]

xxx

\subsection{Controlled Modular Addition}

As shown in figure \ref{fig:mod-add-from-pivot-flip-bars}, a modular addition is three pivot flips.
To add $a$ into a register modulo $r$, perform pivot-flips with the pivot at $r-1-a$, then $r-1$, then $a-1$.
See figure \ref{fig:modular-add} for the circuit.

Be aware that this construction requires that the input register be in range.
If the input is greater than or equal to $r$, the pivot-flips will trash the target register instead of adding the input mod $r$ into it.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/mod-add-from-pivot-flip-bars.png}
  \caption{\em
     A modular offset is three pivot flips.
     One at $r-a$, then one at $r$, then one at $a$.
   }
  \label{fig:modular-add}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/modular-addition.png}
  \caption{\em Reducing modular addition to non-modular offset, pivot-flips, and NOTs.}
  \label{fig:modular-add}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/controlled-modular-addition.png}
  \caption{\em Reducing controlled modular addition to non-modular offset, controlled pivot-flips, and NOTs.}
  \label{fig:modular-add}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/modular-offset.png}
  \caption{\em Reducing modular offset to non-modular offset, controlled pivot-flips, and NOTs.}
  \label{fig:modular-add}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/controlled-modular-offset.png}
  \caption{\em Reducing controlled modular offset to non-modular offset, controlled pivot-flips, and NOTs.}
  \label{fig:modular-add}
\end{figure}

\section{$O(N)$ inline adders}

In \cite{van2004}, a reversible adder with $O(N)$ depth and $O(N)$ size, requiring a single clean ancilla, is described.
If the ancilla is not initialized in the off state, the target register ends up storing $a+b+1$ instead of $a+b$.
For clarity, we review how to elide the ancilla before generalizing the construction to apply to registers of differing size.

The need for the ancilla to be clean can be fixed by inverting (bitwise-NOT-ing) the input and target registers before and after the addition when the ancilla is on.
Because $\lnot x = -x-1$, and the registers are the same size, temporarily inverting the target effectively turns the addition into a subtraction.
Temporarily inverting the input effectively turns the subtraction back into an addition, but also introduces an extra shift by -1, cancelling out the +1 error.

Once the ancilla can be dirty the MSB of the input register can play the role of the ancilla, as shown in figure \ref{fig:inlineadder}.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/inline-adder.png}
  \caption{\em Adder for input and target registers of the same size.
  Requires no ancilla, and uses $O(N)$ gates and depth.
  Based on \cite{van2004, takahashi2005}.}
  \label{fig:inlineadder}
\end{figure}

To perform subtraction instead of addition: run the circuit in reverse, temporarily invert the target register by bordering the circuit with NOT gates on the target, or invert the CSWAP controls.

Because this adder construction uses the input register as workspace, it only works when the input register is the same size as the target register.
When the input register is larger, the high bits of the input register can simply be ignored.
But, when the target register is larger, the workaround is more difficult.
Not only is there insufficient workspace, but the register sizes being different breaks the logical-negation trick that allowed the ancilla to be dirty.

To work around the target register being larger, three fixes are needed.
Each uses a controlled increment or decrement gate, which will be described in the next section.
(Note that, to avoid cyclic dependencies, the increment and decrement constructions will only use the same-register-size adder.)

First, the carry signal must be forwarded into the high part of the target register with a controlled increment instead of with a CNOT.
Second, to free the MSB of the input register for use as the carry signal, we add it into the target ahead of time with a controlled increment.
Finally, when the MSB is on instead of off its role as carry signal causes an extra increment of the target.
This extra increment is cancelled with a controlled decrement.
See the circuit diagram in figure \ref{fig:inline-adder-into-large}.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/inline-adder-into-large.png}
  \caption{\em
      Adder with target larger than source, using no ancilla.
      Uses $O(N)$ gates and depth.
      The increment and decrement gates all have at least one free wire to borrow as a dirty ancilla.}
  \label{fig:inline-adder-into-large}
\end{figure}

\section{$O(N)$ incrementer}

A register can be incremented by subtracting both $x$ and $\neg x = -x-1$ from it, for any $x$.
When $N$ dirty bits are available, $x$ can come from a register defined by those $N$ arbitrary bits, as shown in figure \ref{fig:increment-many-dirty}.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/increment-many-dirty.png}
  \caption{\em Subtracting $x$ and $-x-1$ from a register increments it. Requires $O(N)$ depth, size, and $N$ dirty ancilla.}
  \label{fig:increment-many-dirty}
\end{figure}

To improve from $N$ dirty bits to a single dirty bit, we break the increment operation into two halves.
A high half that is incremented only if all of the bottom bits are on, and a low half that is unconditionally incremented.
The low half can be incremented with the double-subtraction trick by borrowing the high half.
But the double-subtraction trick doesn't work on the high half, because the low half can't be borrowed when it is being used as a control.

To work around not being able to operate on the borrowed bits while using them as a control, start with the double-subtraction trick but change the second subtraction to an addition and conditionally toggle the target bits instead of the input bits.
When the condition isn't satisfied, the addition and subtraction will cancel each other.
When the condition is satisfied, the addition will be inverted into a subtraction, and both the subtraction and addition-turned-subtraction will fire, subtracting the input register from the target register twice.
However, because we are conditioning on the input bits all being on, we know the input must be -1.
Therefore the target was incremented by 2.
To halve the +2 into a +1, prepend a dirty least-significant-bit onto the target register as shown in \ref{fig:increment-odd}.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/increment-odd.png}
  \caption{\em Odd-sized $O(N)$ increment with 1 dirty ancilla.}
  \label{fig:increment-odd}
\end{figure}

Because the adder/subtracter construction used by the incrementer in figure \ref{fig:compact-increment} requires the target and source registers to be the same size, the incrementer construction described so far only works on odd-sized registers.
For even-sized registers, another minor variation on the conditionally-invert-addition-into-subtraction-with-logical-negation technique is used.
That construction is shown in figure \ref{fig:compact-increment-even}.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/increment-even.png}
  \caption{\em Even-sized $O(N)$ increment with 1 dirty ancilla.}
  \label{fig:increment-even}
\end{figure}

As with addition and subtraction, you can turn an increment into a decrement either by reversing the gate order, by bordering the circuit with NOT gates, or by inverting controls.
Also, increment gates can be efficiently controlled by using another invert-input-and-output trick (see figures \ref{fig:controlled-increment-odd} and \ref{fig:controlled-increment-even}).

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/controlled-increment-odd.png}
  \caption{\em Odd-sized $O(N)$ controlled increment with 1 dirty ancilla.}
  \label{fig:controlled-increment-odd}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/controlled-increment-even.png}
  \caption{\em Even-sized $O(N)$ controlled increment with 1 dirty ancilla.}
  \label{fig:controlled-increment-even}
\end{figure}

Note that it is impossible to create a classical incrementer that uses no ancilla.
The parity of the permutation performed by an increment operation is odd, but the parity of the permutation performed by any classical gate that doesn't cover the entire circuit is even.
Quantumly, the parity barrier can be bypassed by using partial rotations (see figure \ref{fig:bootstrap-ancilla}).

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/ancilla-bootstrap.png}
  \caption{\em Bootstrapping a dirty ancilla out of an increment gate using quantum operations.
  The $i^{\text{frac}}$ gate is a ``phase gradient" operation that phases each computational basis state $|v\rangle$ by an amount proportional to $v/2^d$, where $d$ is the size of the register.
  In this case each state is phased by $e^{i \frac{\pi}{2} v/2^d}$.
  The phase gradient is implemented by a column of $Z^{2^{-k}}$ gates, .}
  \label{fig:bootstrap-ancilla}
\end{figure}

\section{$O(N)$ inline comparisons}

We review how to implement comparison operations, which toggle a target bit based on the relationship between two input registers, using addition and subtraction operations.

The is-B-less-than-A comparison is implemented by subtracting $A$ out of $B$, with a target bit appended as $B$'s MSB, then adding $A$ into $B$ without the target bit appended.
See figure \ref{fig:comparison-less}.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/comparison-less.png}
  \caption{\em Toggling a target bit if one register is less than another, without using ancilla.}
  \label{fig:comparison-less}
\end{figure}

The same idea is used for is-B-less-than-or-equal-to-A, except $A+1$ is subtracted and added instead of $A$.
The extra $\pm 1$ can be performed by increment and decrement gates, as shown in figure \ref{fig:comparison-leq}.
When the two registers are the same size, the increment gate can be omitted by inverting the $A$ and $B$ registers during the addition.

$>$ and $\geq$ comparisons are done by performing the corresponding opposite comparison ($\leq$ and $<$ respectively) and inverting the result by toggling the target.

Beware that, when the $B$ register is smaller than the $A$ register, the high bits of $A$ are not accounted for by the subtraction and addition and so the construction breaks.
To fix the problem, swap the roles of A and B (i.e. subtract out of A instead) and use the opposite comparison.
Alternatively, any operations generated by the subtraction or decrement gates that toggle the target bit can be controlled by all of the high bits of $A$ being off and otherwise (if any of the high bits are on) concluding $A>B$ and toggling the result bit as appropriate for the comparison being implemented.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/comparison-leq.png}
  \caption{\em Toggling a target bit if one register is less than or equal to another, without using ancilla.}
  \label{fig:comparison-leq}
\end{figure}

\section{$\text{PivotFlip}$}

\section{$O(N \lg N)$ Modular Addition and Modular Offset} \label{sec:last-circuit}

Although the fact is obfuscated by the number of constructions used, our modular addition construction is performed entirely in terms of normal addition (with various tweaks and tricks and inversions along the way).
See figure \ref{fig:modular-dependencies} for the full dependency graph.

The modular addition uses $O(N \lg N)$ gates because that is the cost of the offset gate in \cite{haner2016}, and we subtract the modulus (typically a compile-time constant) out of a register during one of the pivot-flips.
If both the amount to add {\em and the modulus} are in registers, then the construction only requires $O(N)$ gates.
The total depth is $O(N)$ regardless.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{assets/modular-add-graph.png}
  \caption{\em
      Dependency graph of constructions used to create a classical reversible modular adder / modular offset gates using only 1 dirty bit.}
  \label{fig:modular-dependencies}
\end{figure}

To perform modular multiplication, we use the construction from [[[[ reference ]]], which works by iteratively scaling and adding the target register into an ancilla register.
The ancilla and target register then switch roles, and the old target register is erased by adding in the new target times a  multiplicative inverse.

\section{Factoring with $2n+1$ qubits} \label{sec:applications}

In \cite{haner2016}, H\"{a}ner et al. describe how to perform Shor's algorithm with Toffoli-based circuits using $2n+2$ qubits.
The first $n$ qubits are needed to hold the input.
The second $n$ qubits are used for performing a non-power-of-2 QFT [[[ cite what they used ]]].
One qubit is used for the phase estimation process [[[ cite what they used ]]].
The last qubit is a clean ancilla used by their modular offset operation (to temporarily hold a comparison result).

The modular addition circuit we have presented does not require that last clean ancilla.
Instead, we borrow the phase estimation qubit as a temporary dirty bit.
Our modular offset circuit is a drop-in replacement for the one used by H\"{a}ner et all, with the same asymptotic costs ($O(N \lg N)$ size and $O(N)$ depth).
Doing so creates an implementation of Shor's factoring algorithm that uses $2n+1$ qubits, without increasing the asymptotic size or depth with respect to the $2n+2$ qubit algorithm.

[[[ diagram of Shor's algorithm ]]]

\section{Conclusion} \label{sec:conclusion}

We present linear sized classical reversible circuits for incrementing, modular addition, and modular offset that use only a single dirty bit.
Our modular offset circuit improves the minimum number of qubits used when factoring an $n$-bit integer from $2n+2$ to $2n+1$.

For incrementing, a single dirty ancilla is classically optimal because the permutation defined by an increment operation is odd.
This obstacle does not apply to modular additions, because any odd parities can made even with a useless swap in the range of states above the modulus.
So it may be possible to elide the final dirty ancilla our modular addition construction requires, even when restricted to Toffoli-based circuits.

\bibliographystyle{plain}
\bibliography{citations}

\section{Temporary Test Circuit Links}

Convenience links for myself when checking that things work.

\href
{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B%22Swap%22%2C1%2C1%2C1%2C%22Swap%22%2C%22%E2%80%A2%22%5D%2C%5B1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C%22Swap%22%2C1%2C1%2C%22Swap%22%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C%22Swap%22%2C1%2C%22Swap%22%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C1%2C%22Swap%22%2C%22Swap%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C1%2C%22Swap%22%2C%22Swap%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C%22Swap%22%2C1%2C%22Swap%22%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C%22X%22%5D%2C%5B1%2C%22Swap%22%2C1%2C1%2C%22Swap%22%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C%22X%22%5D%2C%5B%22Swap%22%2C1%2C1%2C1%2C%22Swap%22%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C%22X%22%5D%2C%5B%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%5D%7D}
{5+5-bit inline adder}
(or \href
{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22Counting10%22%5D%2C%5B%22QFT7%22%5D%2C%5B1%2C1%2C1%2C%22QFT7%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B%22Swap%22%2C1%2C1%2C1%2C%22Swap%22%2C%22%E2%80%A2%22%5D%2C%5B1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C%22Swap%22%2C1%2C1%2C%22Swap%22%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C%22Swap%22%2C1%2C%22Swap%22%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C1%2C%22Swap%22%2C%22Swap%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C1%2C%22Swap%22%2C%22Swap%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C%22Swap%22%2C1%2C%22Swap%22%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C1%2C%22X%22%5D%2C%5B1%2C%22Swap%22%2C1%2C1%2C%22Swap%22%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C1%2C%22X%22%5D%2C%5B%22Swap%22%2C1%2C1%2C1%2C%22Swap%22%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22%E2%80%A2%22%2C%22X%22%5D%2C%5B%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22inputA5%22%2C1%2C1%2C1%2C1%2C%22%2B%3DA5%22%5D%2C%5B1%2C1%2C1%2C%22QFT%E2%80%A07%22%5D%2C%5B%22QFT%E2%80%A07%22%5D%2C%5B%22Uncounting10%22%5D%5D%7D}
{with test harness})

\href
{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%5D%2C%5B%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%5D%2C%5B%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%5D%7D}
{4-bit incrementer w/ 4 dirty}

\href
{http://algorithmicassertions.com/quirk#circuit=%7B%22cols%22%3A%5B%5B1%2C1%2C1%2C1%2C1%2C%22%3C%3C4%22%5D%2C%5B1%2C%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22inputA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C%22%2B%3DA4%22%2C1%2C1%2C1%2C%22inputA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22%3E%3E4%22%5D%5D%7D}
{8-bit incrementer w/ 1 dirty}
(or \href
{http://algorithmicassertions.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22Counting9%22%5D%2C%5B%22QFT6%22%5D%2C%5B1%2C1%2C1%2C%22QFT6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22%3C%3C4%22%5D%2C%5B1%2C%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22inputA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C%22%2B%3DA4%22%2C1%2C1%2C1%2C%22inputA4%22%5D%2C%5B%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22%3E%3E4%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22dec8%22%5D%2C%5B1%2C1%2C1%2C%22QFT%E2%80%A06%22%5D%2C%5B%22QFT%E2%80%A06%22%5D%2C%5B%22Uncounting9%22%5D%5D%7D}
{with test harness})

\href
{http://algorithmicassertions.com/quirk#circuit=%7B%22cols%22%3A%5B%5B1%2C1%2C1%2C1%2C1%2C%22%3C%3C5%22%5D%2C%5B%22inputA5%22%2C1%2C1%2C1%2C1%2C%22-%3DA5%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22inputA5%22%2C1%2C1%2C1%2C1%2C%22%2B%3DA5%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22-%3DA5%22%2C1%2C1%2C1%2C1%2C%22inputA5%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22-%3DA5%22%2C1%2C1%2C1%2C1%2C%22inputA5%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22%3E%3E5%22%5D%5D%7D}
{9-bit incrementer w/ 1 dirty}
(or \href
{http://algorithmicassertions.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22Counting10%22%5D%2C%5B%22QFT7%22%5D%2C%5B1%2C1%2C1%2C%22QFT7%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22%3C%3C5%22%5D%2C%5B%22inputA5%22%2C1%2C1%2C1%2C1%2C%22-%3DA5%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22inputA5%22%2C1%2C1%2C1%2C1%2C%22%2B%3DA5%22%5D%2C%5B%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22-%3DA5%22%2C1%2C1%2C1%2C1%2C%22inputA5%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22-%3DA5%22%2C1%2C1%2C1%2C1%2C%22inputA5%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C%22%3E%3E5%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22dec9%22%5D%2C%5B1%2C1%2C1%2C%22QFT%E2%80%A07%22%5D%2C%5B%22QFT%E2%80%A07%22%5D%2C%5B%22Uncounting10%22%5D%5D%7D}
{with test harness})

\href
{http://algorithmicassertions.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22Z%5E%E2%85%9F%E2%82%81%E2%82%86%22%2C%22Z%5E%E2%85%9F%E2%82%81%E2%82%86%22%2C%22Z%5E%E2%85%9B%22%2C%22Z%5E%C2%BC%22%2C%22H%22%5D%2C%5B%22%E2%80%A2%22%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C%22Z%5E-%C2%BC%22%5D%2C%5B1%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C%22Z%5E%C2%BC%22%5D%2C%5B%22%E2%80%A2%22%2C1%2C1%2C1%2C%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C%22Z%5E-%C2%BC%22%5D%2C%5B1%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22%E2%80%A2%22%2C%22X%22%5D%2C%5B%22inc4%22%2C1%2C1%2C1%2C%22Z%5E%C2%BC%22%5D%2C%5B1%2C%22Z%5E-%E2%85%9F%E2%82%81%E2%82%86%22%2C%22Z%5E-%E2%85%9B%22%2C%22Z%5E-%C2%BC%22%2C%22H%22%5D%5D%7D}
{Quantum incrementer ancilla bootstrap}

NOTE: have to open these next ones against local version, because online version doesn't have the comparison gates

\href
{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3EB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3EB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%5D%7D}
{4+4-Bit $<$ PivotFlip}
(or \href
{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B1%2C1%2C1%2C1%2C%22H%22%2C%22H%22%2C%22H%22%2C%22H%22%5D%2C%5B%22Counting4%22%2C1%2C1%2C1%2C%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%2C1%2C1%2C1%2C%22X%5Et%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22%3C%3C5%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3EB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3EB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22%3E%3E5%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22X%5E-t%22%5D%2C%5B%22Chance4%22%2C1%2C1%2C1%2C%22Amps8%22%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22Chance%22%5D%5D%7D}
{with test harness})

\href
{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3E%3DB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22dec4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3E%3DB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22dec4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%5D%7D}
{4+4-Bit $\leq$ PivotFlip}
(or \href
{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B1%2C1%2C1%2C1%2C%22H%22%2C%22H%22%2C%22H%22%2C%22H%22%5D%2C%5B%22Counting4%22%2C1%2C1%2C1%2C%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%2C1%2C1%2C1%2C%22X%5Et%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22%3C%3C5%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3E%3DB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22dec4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22inputB4%22%2C1%2C1%2C1%2C%22%5EA%3E%3DB%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22-%3DA4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22dec4%22%2C1%2C1%2C1%2C%22%E2%80%A2%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%2C%22%E2%80%A2%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22%3E%3E5%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22X%5E-t%22%5D%2C%5B%22Chance4%22%2C1%2C1%2C1%2C%22Amps8%22%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22Chance%22%5D%5D%7D}
{with test harness})

\href
{http://algassert.com/quirk#circuit=%7B%22cols%22%3A%5B%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22inc4%22%5D%2C%5B%22X%22%2C1%2C1%2C%22X%22%2C%22Counting4%22%2C1%2C1%2C1%2C%22QFT4%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%5D%2C%5B%22Chance4%22%2C1%2C1%2C1%2C%22Chance4%22%2C1%2C1%2C1%2C%22Chance4%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B1%2C1%2C1%2C1%2C%22inc4%22%5D%2C%5B1%2C1%2C1%2C1%2C%22inputA4%22%2C1%2C1%2C1%2C%22Flip%3CA4%22%5D%2C%5B1%2C1%2C1%2C1%2C%22dec4%22%5D%2C%5B1%2C1%2C1%2C1%2C%22X%22%2C%22X%22%2C%22X%22%2C%22X%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22Flip%3CA4%22%5D%2C%5B%22inputA4%22%2C1%2C1%2C1%2C%22%2B%3DA4%22%5D%2C%5B1%2C1%2C1%2C1%2C%22inputA4%22%2C1%2C1%2C1%2C%22Flip%3CA4%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22%E2%80%A6%22%5D%2C%5B%22Chance4%22%2C1%2C1%2C1%2C%22Chance4%22%2C1%2C1%2C1%2C%22Amps8%22%5D%2C%5B%5D%2C%5B%5D%2C%5B%5D%2C%5B%5D%2C%5B%5D%2C%5B%5D%2C%5B%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22%7C0%E2%9F%A9%E2%9F%A80%7C%22%5D%2C%5B%22inputB4%22%2C1%2C1%2C1%2C%22inputA4%22%2C1%2C1%2C1%2C%22%5EA%3E%3DB%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22Chance%22%5D%2C%5B1%2C1%2C1%2C1%2C1%2C1%2C1%2C1%2C%22~a82g%22%5D%5D%2C%22gates%22%3A%5B%7B%22id%22%3A%22~a82g%22%2C%22name%22%3A%22IGNORE%22%2C%22matrix%22%3A%22%7B%7B1%2C0%7D%2C%7B0%2C1%7D%7D%22%7D%5D%7D}
{Modular Addition with test Harness}

\end{document}